---
title: "STAT HW04"
author: "Xinmiao Wang"
date: "`r format(Sys.Date())`"
output: github_document
---

#Navigation 

* The main repo for homework: [here](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao) 

* Requirement for Homework 04: click [here](http://stat545.com/hw04_tidy-data-joins.html)

* hw04 folder: [here](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/tree/master/hw04).

* Files inside hw04:

  1. [README.md](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw04/README.md)
  2. [hw04_Gapminder.Rmd](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw04/hw04_Gapminder.Rmd)
  3. [hw04_Gapminder.md](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw04/hw04_Gapminder.md)


# Induction

The goal of this homework is to solidify your data wrangling skills by working some realistic problems in the grey area between data aggregation and data reshaping.

In this Homework, we still work on Gapminder dataset, and are going to exploring the dataset by reshaping it. So, We need to load the packages, `gapminder`, `tidyverse` and `devtools`.

# Load Package

Install `gapminder` from CRAN:
```{r eval=FALSE}
install.packages("gapminder")
```

Install `tidyverse` from CRAN:
```{r eval=FALSE}
install.packages("tidyverse")
```

Install `devtools` from CRAN:
```{r eval=FALSE}
install.packages("devtools")
```

Here, we load packages that we gonna use later.

```{r load_library, warning=F}
#load packages
library(tidyverse)
library(gapminder)
library(devtools)
library(ggthemes)
library(countrycode)
library(geonames)
```


# General data reshaping and relationship to aggregation

## Activity 2

*Make a tibble with one row per year and columns for life expectancy for two or more countries.

  - Use knitr::kable() to make this table look pretty in your rendered homework.
  - Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.
  
```{r reshaping}
gapminder.part <- gapminder %>% 
  filter(continent == "Americas") %>% 
  select(country, lifeExp, year, gdpPercap)

dat1 <- gapminder.part %>% 
  select(country, lifeExp, year) %>% 
  spread(country, lifeExp) %>% 
  as.tbl()

knitr::kable(dat1)

ggplot(dat1, aes(x=Canada, y=get("United States")))+
  geom_point()+
  labs(x="Canada", y="United States")+
  theme_calc()+
  ggtitle("The Plot of Life Expentancy in US verse it in Canada")

```


# Join, merge, look up

## Activity 1

* Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a  dplyr join function and make some observations about the process and result. Explore the different types of joins. Examples of a second data frame you could build:

  - One row per country, a country variable and one or more variables with extra info, such as language spoken, NATO membership, national animal, or capitol city. If you really want to be helpful, you could attempt to make a pull request to resolve this issue, where I would like to bring ISO country codes into the gapminder package.
  - One row per continent, a continent variable and one or more variables with extra info, such as northern versus southern hemisphere.


I am going to use the data in `countrycode` to creat a new dataset which include the information of countries in North America. 

```{r echo=F}
options(geonamesUsername="xinmiaow")
```

```{r creat_dataset}
countryinfo <- GNcountryInfo() %>% 
  tbl_df %>%
  filter(continent=="NA") %>% 
  select(country=countryName, iso=isoAlpha3, capital=capital, lang=languages) %>%
  mutate(country=as.character(country), iso=as.character(iso), capital=as.character(capital), lang=as.character(lang))

knitr::kable(countryinfo)

```

```{r join_merge}
gapminder.part %>% 
  left_join(countryinfo, by="country") %>% 
  knitr::kable()

```



