---
title: "STAT HW04"
author: "Xinmiao Wang"
date: "`r format(Sys.Date())`"
output: github_document
---

#Navigation 

* The main repo for homework: [here](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao) 

* Requirement for Homework 04: click [here](http://stat545.com/hw04_tidy-data-joins.html)

* hw04 folder: [here](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/tree/master/hw04).

* Files inside hw04:

  1. [README.md](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw04/README.md)
  2. [hw04_Gapminder.Rmd](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw04/hw04_Gapminder.Rmd)
  3. [hw04_Gapminder.md](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw04/hw04_Gapminder.md)


# Induction

The goal of this homework is to solidify your data wrangling skills by working some realistic problems in the grey area between data aggregation and data reshaping.

In this Homework, we still work on Gapminder dataset


# Load Package

Install `gapminder` from CRAN:
```{r eval=FALSE}
install.packages("gapminder")
```

Install `tidyverse` from CRAN:
```{r eval=FALSE}
install.packages("tidyverse")
```

Install `devtools` from CRAN:
```{r eval=FALSE}
install.packages("devtools")
```

Here, we load packages that we gonna use later.

```{r load_library, warning=F}
#load packages
library(tidyverse)
library(gapminder)
library(devtools)
library(ggthemes)
```


# General data reshaping and relationship to aggregation

## Activity 2

*Make a tibble with one row per year and columns for life expectancy for two or more countries.

  - Use knitr::kable() to make this table look pretty in your rendered homework.
  - Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.
  
```{r reshaping}

```

