---
title: "STAT545 HW02"
author: "Xinmiao Wang"
date: "`r format(Sys.Date())`"
output: github_document
---

# Bring Rectangular Data in

In this module, we intend to explore Gapminder data and practice the functions in dplyr, which can be loaded from gapminder package and tidyverse package in R. Please make it sure that those package have been installed before we load them.

Install `gapminder` from CRAN:
```{r eval=FALSE}
install.packages("gapminder")
```

Install `tidyverse` from CRAN:
```{r eval=FALSE}
install.packages("tidyverse")
```

Here, we load those two packages.

```{r load_library, warning=F}
#load packages
library(gapminder)
library(tidyverse)
```


# Smell Test the Data

* Q: Is it a data.frame, a matrix, a vector, a list?

```{r type_data}
gapminder
str(gapminder)
```

  A: The data is a tibble, because we load the `tidyverse` package. It is also a data.frame by the classes shown in `str(gapminder)`.


* Q: What¡¯s its class?

```{r class_data}
class(gapminder)
```
 
  A: The class of gapminder includes `r class(gapminder)`.


* Q: How many variables/columns?

```{r ncol_data} 
ncol(gapminder)
```

  A: There are `r ncol(gapminder)` variables.


* Q: How many rows/observations?

```{r nrow_data}
nrow(gapminder)
```

  A: A: There are `r nrow(gapminder)` observations.
  
  
* Q: Can you get these facts about ¡°extent¡± or ¡°size¡± in more than one way? Can you imagine different functions being useful in different contexts? 

```{r other_extent}
dim(gapminder) # number of observations and number of variables
length(gapminder) # number of variables
```

  A: We can also use functions `dim()` to check the basic extent, and use `length()` to check the number of variables. In addition, we can also get the extent by using `str()`.
  
  
* Q: What data type is each variable?

```{r type_var}
attach(gapminder)
a <- rbind(names(gapminder), c(typeof(country), typeof(continent), typeof(year), typeof(lifeExp), typeof(pop), typeof(gdpPercap)))
as.data.frame(a, row.names=c("Variables", "Type"))

```

  A: The table above lists the types of all variabls. 
     + Integer: country, contient, year, and pop
     + Double: lifeExp and gdpPercap


# Explore individual variables

* Categorical Variable:  

* quantitative Variable