---
title: "STAT545 HW10"
author: "Xinmiao Wang"
date: "December 3, 2017"
output: github_document
---

# Navigation 

* The main repo for homework: [here](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao) 

* Requirement for Homework 10: click [here](http://stat545.com/hw10_data-from-web.html)

* hw10 folder: [here](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/tree/master/hw10).

* Files inside hw10:

  1. [README.md](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw10/README.md)
  2. [hw10_IMDb.Rmd](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw10/hw10_IMDb.Rmd)
  3. [hw10_IMDb.md](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw10/hw10_IMDb.md)
  

# Introduction 

In this assignment, we practice collecting data from website. Below, I collect two datasets from different sources. 

1. `topmovie`: data from [IMDb](http://www.imdb.com/?ref_=nv_home), contains 250 movies which are the top of 250 rated movies. The variables are title, rank, year and rate.

2. `movie.info`: data from [OMDb API](http://www.omdbapi.com/), contains the information of movies whose title and year can match to the source data. The variables include title, rank, runtime, country and genre, 

After that, I merge the two dataset according to the rank and do some basic explanatory analysis for some variables.

# Load Package

```{r warning=FALSE}
library(tidyverse)
library(magrittr)
library(purrr)
library(glue)
library(stringr)

library(xml2)
library(rvest)

library(httr)

library(ggthemes)
```

# Read Data

## Read the page of Top Rated Movies: Scrape Data

I collect the data from [IMDb:Top Rated Movies](http://www.imdb.com/chart/top?ref_=nv_mv_250_6) 
```{r}
url_title <- "http://www.imdb.com/chart/top?ref_=nv_mv_250_6"
page_title <- read_html(url_title)
```

### Read the title of Top Rated Movies
```{r}
title <- page_title %>% 
  html_nodes(css = ".titleColumn") %>% 
  html_text()

head(title, 5)

```

As we can see above, the format of movies' titles are really messy. Each record contains the order of the movie's rank, the title of the movie and also the year in which the movie was produced. Here, I am going to write a function so that we can extract the titles and year easily.

```{r}
get.title <- function(x){
  x.tmp <- str_extract_all(x, boundary(("word")))[[1]]
  n <- length(x.tmp)
  title <- str_c(x.tmp[-c(1, n)], collapse = " ")
  return(title)
}

get.year <- function(x){
  x.tmp <- str_extract_all(x, boundary(("word")))[[1]]
  n <- length(x.tmp)
  year <- as.integer(x.tmp[n])
  return(year)
}

get.rank <- function(x){
  x.tmp <- str_extract_all(x, boundary(("word")))[[1]]
  rank <- as.integer(x.tmp[1])
  return(rank)
}
```

I read the movies' titles and year by using the functions `get.title` and `get.year` above.
```{r}
mytitle <- map_chr(title, get.title)

head(mytitle, 5)

myyear <- map_int(title, get.year)

head(myyear, 5)

myrank <- map_int(title, get.rank)

head(myrank, 5)

```

### Read the Rates of Movies
```{r}
rate <- page_title %>% 
  html_nodes(css = ".imdbRating") %>% 
  html_text()

head(rate, 5)

```

Similar to the last part, the format of rate is also not easy to read. Here, I define a function so that we can extract the rate of each movie as a double.
```{r}
get.rate <- function(x){
  x.tmp <- str_extract_all(x, boundary(("word")))[[1]]
  n <- length(x.tmp)
  rate <- as.double(x.tmp[n])
  return(rate)
}
```

Now, let's extract the rate. We can see it can be used and analyzed easily.
```{r}
myrate <- map_dbl(rate, get.rate)

head(myrate, 5)
```

### Combine Variables
After collecting all the information in this page, I am going to combine these three variables as a dataset.

```{r}
topmovie <- data_frame(title=mytitle, rank=myrank, year=myyear, rate=myrate)

head(topmovie, 5)
```


## Read data for Each Movie: APIs
At first, I tried to extract the information of each movie from their own page. However, it is difficult to find the corresponding tag for all information I want to collect. Hence, I am going to collect the information of each movie from OMDb API.

### Define Function: get.info
Here, I define a function `get.info()` to extract the information for a movie by the title and year. If the response in `moive_content` is true, there is the information for the movie, so we extract the information; otherwise, all the corresponding information will be returned as NA.

```{r}
get.info <- function(title, year, rank.t){
  title.dat <- str_c(str_split(title, " ", simplify = TRUE), collapse = "+")
  year.dat <- as.character(year)
  query_string <- glue("http://www.omdbapi.com/?t={title.dat}&y={year.dat}&apikey=809490ed")
  movie_result <- GET(query_string)
  movie_content <- content(movie_result)
  if (as.logical(movie_content$Response)){
    title.tmp <- movie_content$Title
    runtime.tmp <- as.integer(str_sub(movie_content$Runtime, 1, 3))
    country.tmp <- str_split(movie_content$Country, ",", simplify = TRUE)[1]
    language.tmp <- str_split(movie_content$Language, ",", simplify = TRUE)[1]
    genre.c <- movie_content$Genre
    genre.tmp <- str_split(genre.c, ",", simplify = TRUE)[1]
  } else {
    title.tmp <- title
    runtime.tmp <- NA
    country.tmp <- NA
    language.tmp <- NA
    genre.tmp <- NA
  }
  output <- data_frame(title=title.tmp, rank=rank.t, runtime=runtime.tmp, country=country.tmp, language=language.tmp, genre=genre.tmp)
  return(output)
}
```

### Extract Infomation 
```{r}
movie.info <- pmap_df(list(topmovie$title, topmovie$year, topmovie$rank), get.info)

head(movie.info, 5)

# Check the number of missing values
sum(is.na(movie.info$runtime))
```

## Merge Dataset
To combine the information from different source, I will merge the two datatsets so that we can use the information later.
```{r}
movie.dat.tmp <- topmovie %>% 
  left_join(movie.info, by="rank")

knitr::kable(head(movie.dat.tmp, 10))

```

## Remove Title
Because I merge the two dataset by `rank`, there are two variables about title in our final dataset. As we can see above, there are some differences between the title from two dataset. It is because of the way I clean the title collected from IMDb website. The tile from OMDb are supposed to be a right format. Hence, I am going to remove the title.x
```{r}
movie.dat <- movie.dat.tmp %>% 
  mutate(title.x = title.y) %>% 
  select(-title.y) %>% 
  rename(title=title.x)

knitr::kable(head(movie.dat, 10))
```

## Check the merged dataset
The number of observations are supposed to be 250;
The number of missing values are supposed to be 41
```{r}
# The number of observations (250)
nrow(movie.dat)
# The number of missing values (41)
sum(is.na(movie.dat$runtime)) 
```


# Dataset

`movie. dat` collects top 250 rated movies from IMDb and contains the information about each movie.

In the dataset, there are `r nrow(movie.dat)` observations with `r ncol(movie.dat)` variables,

**Catergorical variables**
   - title: The title of each movie
   - country: the country where the movie were produced. 
   - language: the language of the movie
   - genre: the genre of the movie

**Quantitative variables**
   - rank: the rank of the movie
   - year: when the movie were produced
   - rate: the rate of the movie
   - runtime: the running time of the movie
   
```{r}
str(movie.dat)
```

# Explanatory Analysis

## Country 

Most of top rated movie were produced in USA.
```{r warning=FALSE}
movie.dat %>% 
  group_by(country) %>% 
  summarise(count = length(title)) %>% 
  t() %>% 
  knitr::kable()

```

```{r}
movie.dat %>%
  group_by(country) %>% 
  filter(length(title)>=3) %>% 
  ggplot(aes(x=country))+
  geom_bar(aes(color=country, fill=country))+
  theme_calc()+
  ggtitle("The Bar Chart of Country (count>=3)")
```


## Genre

Most of top rated movies are Drama.
```{r warning=FALSE}
movie.dat %>% 
  group_by(genre) %>% 
  summarise(count = length(title)) %>% 
  t() %>% 
  knitr::kable()

```

```{r}
movie.dat %>%
  group_by(genre) %>% 
  filter(length(title)>1) %>% 
  ggplot(aes(x=genre))+
  geom_bar(aes(color=genre, fill=genre))+
  theme_calc()+
  ggtitle("The Bar Chart of Genre (count>1)")
```


## Runtime

Most of top rated movies lasts for around 130 minutes.
```{r}
knitr::kable(t(summary(movie.dat$runtime)))
```

```{r warning=FALSE}
movie.dat %>% 
  ggplot(aes(x=runtime))+
  geom_histogram(binwidth = 5,col="red", aes(fill=..count..))+
  scale_fill_gradient("count", low = "green", high = "red")+
  theme_calc()+
  ggtitle("The Histogram of Running Time")
```



## Runtime verse Year

There is no exact pattern of the running time over year. Similar to the histogram, Most of top rated movies lasts for between 100 minutes and 150 minutes

```{r}
movie.dat %>% 
  ggplot(aes(x=year, y=runtime, group=year))+
  geom_point(na.rm = TRUE)+
  theme_calc()+
  ggtitle("The Plot of Running Time over year")
```


## Rate

The minimum rate among these top-rated movies is 8.
```{r}
knitr::kable(t(summary(movie.dat$rate)))
```


## Rate verse Year

```{r}
movie.dat %>% 
  ggplot(aes(x=year, y=rate, group=year))+
  geom_point(aes(color=genre), na.rm = TRUE)+
  theme_calc()+
  ggtitle("The Plot of Rate over year")
```

# Reference

- [IMDb: Top Rated Movies](http://www.imdb.com/chart/top?ref_=nv_mv_250_6)

- [OMDb API](http://www.omdbapi.com/)

- [scraping-data-without-an-api](https://github.com/ropensci/user2016-tutorial/blob/master/03-scraping-data-without-an-api.pdf)
