---
title: "STAT545 HW06"
author: "Xinmiao Wang"
date: "November 4, 2017"
output: github_document
---

# Navigation 

* The main repo for homework: [here](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao) 

* Requirement for Homework 05: click [here](http://stat545.com/hw06_data-wrangling-conclusion.html)

* hw05 folder: [here](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/tree/master/hw06).

* Files inside hw06:

  1. [README.md](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw06/README.md)
  2. [hw06_xw.Rmd](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw06/hw06_xw.Rmd)
  3. [hw06_xw.md](https://github.com/xinmiaow/STAT545-hw-Wang-Xinmiao/blob/master/hw06/hw06_xw.md)
  

# Introduction 

## Load package

```{r load_package}
library(tidyverse)
library(stringr)
library(gapminder)

```


# Character data

## 14.2.5 Exercises

###  1. In code that doesn¡¯t use stringr, you¡¯ll often see paste() and paste0(). What¡¯s the difference between the two functions? What stringr function are they equivalent to? How do the functions differ in their handling of NA?

**ANS:**
1. Differences

   * `paste(...)` separate each inputs with a space automatically; `paste0(...)` can not do that, but it works more efficient when you want to combine inputs as one character.
   * In `paste(...)`, you can control how the inputs are separated by arguement `sep=...`; `paste0(...)` can not do that

2. Equivalent function in Stringr

   * `paste(...)` has the same result as `str_c(..., sep=" ")`
   * `paste0(...)` has the same result as `str_c(...)`

3. Handling of `NA`
  * `paste(...)` and `paste0(...)` treat `NA` as a character
  * `str_c(...)` will return `NA` if any input is `NA` inside of this function
  * You can use `str_replace_na()` inside the function `str_c(...)` so that the input `NA` can be treated as a character
  
```{r p1_q1}
string <- c("This is my", "repo for HW06")

# Combine Multiple inputs
paste(string[1], string[2])

str_c(string[1], string[2],  sep = " ")

paste0(string[1], string[2])

str_c(string[1], string[2])

# Combine Multiple elemnts in a vectorized input
paste(string, collapse = " ")

paste0(string, collapse = " ")

str_c(string, collapse = " ")

# with input NA

string2 <- c("This is my", NA, "?")

paste(string2, collapse = " ")

paste0(string2, collapse = " ")

str_c(string2, collapse = " ")

str_c(str_replace_na(string2), collapse = " ")

```


### 2. In your own words, describe the difference between the sep and collapse arguments to str_c().

```{r p1_q2}
str_c(string, collapse=" ")


```


### 3. Use str_length() and str_sub() to extract the middle character from a string. What will you do if the string has an even number of characters?

### 4. What does str_wrap() do? When might you want to use it?

### 5. What does str_trim() do? What¡¯s the opposite of str_trim()?

### 6. Write a function that turns (e.g.) a vector c("a", "b", "c") into the string a, b, and c. Think carefully about what it should do if given a vector of length 0, 1, or 2.




# Writing functions


